Reproducible Research Course Project 1
======================================

#Loading the dataset in Rstudio
```{r,echo=TRUE}
activity <- read.csv("activity.csv")
summary(activity)
```
#Coverting date from character to date
```{r,echo=TRUE}
activity$date <- as.Date(activity$date,format="%Y-%m-%d")
summary(activity)
```

#calculating number of steps each day
```{r,echo=TRUE}
stepsperday <- aggregate(x=activity$steps,by=list(activity$date),FUN=sum,na.rm=TRUE)
colnames(stepsperday) <- c("date","steps")
```
#Plotting the histogram for number of steps each day 
```{r,echo=TRUE}
hist(x=stepsperday$steps,xlab="steps",main="Number of steps each day",
     breaks = seq(0,25000,by=2500),col="red")
```
#Calculating mean and median of number of steps each day
```{r,echo=TRUE}
mean(stepsperday$steps)
median(stepsperday$steps)
```
#Calculating average of number of steps taken
```{r,echo=TRUE}
avgsteps <- aggregate(x=activity$steps,by=list(activity$interval),FUN=mean,na.rm=TRUE)
colnames(avgsteps) <- c("interval","steps")
```
#Plotting time series plot for average number of steps taken 
time series plot is plotted by type="l"
```{r,echo=TRUE}
plot(x=avgsteps$interval,y=avgsteps$steps,type="l",xlab= "5- mintues interval",
     ylab= "average number of steps taken",main="")
```

#Which 5-minute interval, on average across all the days in the dataset, contains maximum number of steps ?
```{r,echo=TRUE}
intervalwithmaxsteps <- avgsteps[which.max(avgsteps$steps),]$interval
intervalwithmaxsteps
```
#Code to describe and a stratergy for imputing missing data
```{r,echo=TRUE}
sum(is.na(activity$steps))
imputedsteps <- avgsteps$steps[match(activity$interval,avgsteps$interval)]
activityimputed <- transform(activity, steps = ifelse(is.na(activity$steps),
                                  yes = imputedsteps, no = activity$steps))
totalstepsimputed <- aggregate(x=activityimputed$steps,by=list(activityimputed$date),
                              FUN=sum)
names(totalstepsimputed) <- c("date", "daily_steps")
```
#Histogram of the total number of steps taken each day after missing values are imputed
```{r,echo=TRUE}
hist(x=totalstepsimputed$daily_steps,xlab = "Total steps per day",
     main="Total number of steps taken each day",breaks = seq(0,25000,by=2500),
     col="blue")
```

#Classifying the data into weekdays and weekends
```{r,echo=TRUE}
activity$day <- weekdays(activity$date)
head(activity)
activity$day <- ifelse(activity$day=='Saturday' | activity$day=='Sunday', 'weekend','weekday')
```
#Plotting average number of steps taken 5-minute interval across weekdays and weekends using ggplot
```{r,echo=TRUE}
stepsbydate <- aggregate(activity$steps ~ activity$interval + activity$day,FUN = mean,na.rm=TRUE)
names(stepsbydate) <- c("interval","day","steps") 
library(ggplot2)
p <- ggplot(stepsbydate,aes(x=interval,y=steps,color=day))
p + geom_line() + facet_grid(day~.) + xlab("interval") + ylab("Average number of steps") +
  ggtitle("Average daily steps by the type of date")
```